An Innovative Storage Stack Addressing Extreme Scale Platforms and Big Data Applications

### 2014-05-01 paper.tex
filled out DAOS and VOSD sections and added a demonstration section.

The rest of the paper is organized as follows. A brief overview of related work
is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array. This will be discussed in the contect of the HDF-5 based
example library used for the functionality demonstration. Section~\ref{sec:iof}
briefly discusses the motivation and proposal for the IO forwarding layer.
Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
functionality it offers. This will detail the pieces of the layer that are
potentially optional and mention the cross-cutting features discussed in a
later, cross-cutting section. Section~\ref{sec:daos} discusses how the
persistent storage array interface itself functions. As with the IOD layer, the
disucssion of cross-cutting features will be mentioned, but disucssed more
fully in the cross-cutting section. The VOSD layer is disucssed in
Section~\ref{sec:vosd}. In particular, the mapping between the DAOS and VOSD
layers are explored as it pertains to the physical storage. Next is an
exploration of cross-cutting features like transactions and metadata management
in Section~\ref{sec:broader}. Since these and other features are spread across
multiple layers, it makes more sense to discuss them independently once an
understanding of the overall structure has been presented.  The paper is
concluded in Section~\ref{sec:conclusion} with a summary of the broad issues
covered in the paper.

@@

The IOD VOL plugin serves as the bridge between HDF5 and the IOD API. It
is also shown in Figure \ref{vol-arch}. The application calls the HDF5
library while running on the system's compute nodes. Using the VOL
architecture, the IOD VOL plugin uses a function shipper (RPC library)
to forward the VOL calls to a server component running on the I/O nodes
(IONs). At that point, the VOL calls are translated into I/O Dispatcher
(IOD) API calls and executed at the IONs.

@@

This section will briefly describe the limited functionality of the IO
Forwarding layer and why it was included in the design.  It will not exceed
probably half of a column.

@@

In current proposals, the stack is managed by middleware that sits between the
application and the parallel file system. Since most of the applications assume
a POSIX interface to storage, existing middleware either transparently handle
the I/O operations, appearing as regular POSIX calls to applications; or modify
the I/O API as little as possible in order to minimize the impact on production
codes.

As we move towards the exascale goal, this way of interfacing with storage
becomes more of an obstacle. Many features provided by the stack are hidden by
the middleware/POSIX layers. If these were visible to the applications, and if
the applications were able to control how the underlying layers behave, they
would benefit greatly since the domain knowledge that is available at the
application level could be used to execute storage operations efficiently. One
more implication of the 25-year-old POSIX API is in cumbersome middleware,
since metadata has to be part of the same uni-dimensional byte array being
stored.

The Fast Forward Storage I/O project is aimed at merging the features of
existing middleware into a next-generation parallel file system stack. This new
storage API termed the I/O dispatcher (IOD) is object-based, transactional,
asynchronous and I/O-staging-aware (Figure~\ref{fig:exa-stack} (right)).

@@

The prototype stack, from a top-down point of view, has the following
components:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Arbitrarily Connected Graphs (ACG).
\item
  HDF5 extensions.
\item
  I/O Dispatcher (IOD).
\item
  Distributed Asynchronous Object Storage (DAOS).
\item
  Versioning Object Storage Device (VOSD).
\end{enumerate}

Applications running on the compute nodes will be written in Python scripts or
in C++ using the HDF5 APIs. These applications will execute analysis on
graph-based data (ACG). The point is to demonstrate both the HPC and BigData
use cases of the Exascale architecture. The data structures will be stored in
HDF5, for which new API extensions will be implemented. These new API calls
will expose the transactional, asynchronous and function-forwarding semantics
of the underlying stack.

@@

We give a high-level description of each of layer.

@@

One of the bigger concerns is the observation that the original data staging
proposals all used compute nodes while the newer proposals seek not only to
make them a fixed portion of the IO stack, but also shared across all machine
users. The PreDatA~\cite{zheng:2010:predata} paper in particular examines the
potential costs and advantages of where to place operators similar to the IOD
proposed function shipping. There are two key takeaways from PreDatA. First,
placement matters.  Depending on the communication intensity vs. computation
intensity, where along the IO path to place the operation can matter
significantly. Second, and more importantly, the amount of time spent
processing for the operators was stretched to the point where it consumed
nearly all of the time between IO operations. The given ratios of compute
processes to staging process examined is representative for future extreme
scale platforms. If anything, the ratios offer more staging processes than IOD
processes would be available.

In the case of the written IOD design, it describes a fixed-sized staging area
that is partitioned on a per-application basis.  This is unlikely to be useful
because of the limited compute and communication capacity to spare to perform
these operations at a bottleneck in the IO path.  The use of a separate data
staging area intentionally separate from the IO path allows using operators on
limited resources leaving the IO path clear for strictly data movement. A
nuance of this design is discussed in the Design Philosophy below.

By concentrating the Burst Buffers and function shipping into the storage
stack, three problems arise.  First, the amount of network bandwidth, IO
bandwidth, and compute power consumed for example operations from a single
application is likely to completely monopolize the IOD processes. Second, if
space and time partitioning is used instead, the functionality risks being too
small to be useful. Third, the long-term hardware performance advantage for
SSDs is questionable. Recent studies have shown that the erase-before-write
and interference between reading and writing with flash-based SSDs can cause
severe performance problems~\cite{skourtis:2013:ssd-performance}. The inclusion
of an optional use SSD layer in the new Trinity machine at Los Alamos will
offer a test bed to determine how likely these observed problems would affect a
production extreme scale platform.

Given these features, the optionality and even incorporation of burst buffers
in the current design should be carefully considered. Much of the advanced, key
functionality proposed as they are currently designed ultimately relies on the
existence of burst buffers to work. Further thought about how to have an IOD
layer both with and without a burst buffer is required before they can be
considered optional. As the design stands today, they are a required part of
the IOD layer for proper functioning. Unfortunately, it is not clear that they
can address the performance concerns they are intended to cover.

@@

\section{DAOS Layer}\label{sec:daos}
The Distributed Asynchronous Object Storage layer serves as the traditional
file-system interface layer for the storage devices.

\section{VOSD Layer}\label{sec:vosd}
The Versioning Object Storage Device layer operates as the layer for each
persistent storage device used to support the parallel storage array.

@@

At a broader level, there are some concerns that were partially clarified
through conversations with the team.  Consider a shared file system across an
HPC data center. The current design maintains the metadata in its own
container. Since copying data from the IOD layer to the DAOS layer requires an
explicit persist call, how and when synchronizing the metadata across the
layers and potentially across machines occurs is undefined. Delaying
synchronization until an explicit persist is called will reduce the update
frequency, but delays the data visibility on other platforms. Ideally, the
metadata object would need to be automatically persisted every time a container
transaction is persisted to the DAOS layer.

The implication of this is that every transaction persist is double operation
to account for the metadata persist. More importantly, the IOD-layer version of
the metadata container may contain readable transactions that have not been
persisted to the DAOS layer. How to handle this inconsistency between the two
layers still needs to be explored.

A point of confusion rather than a potential design challenge is the change in
definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
container is a collection of objects. For the DAOS layer, a container is a
collection of objects across a set of shards. For the IOD an object may be a
shard of a global array.  For DAOS, a shard can host a set of DAOS objects.
Having the same names with locally correct, globally conflicting definitions
serves to confuse how the system should work.

@@

The transaction mechanism manifests in two forms. At the IOD layer, they are
called transactions and are used to judge whether or not a set of distributed,
asynchronous modifications across a set of related objects is complete or not.
It is also used to control access by treating the transaction ID of committed
transaction as a version identifier.  At the DAOS layer, they are called epochs
and represent persisted (durable) transactions from the IOD layer. Each of
these offers different functionality, but are connected as is explained below.
How these differ from the \DDT approach is also explored.  While IOD's and
\DDTns's transactions are seemingly very different, they use a similar
high-level design, but very different implementation, to solve the same
problem.

@@

Based on the lessons from the \DDT metadata
service~\cite{lofstead:2012:txn-metadata} construction and the prior
experiments with LWFS, having a completely separate metadata service is
workable. Rather than making it a bottleneck in the IO path, it is another
service that users must interact with if they need those services.  Users can
manage everything by maintaining the metadata including the list of objects
themselves. However, there are drawbacks to this 100\% client-side approach.

With a client-side only approach, there is a serious risk of the metadata
service and the object store getting out of sync.  While having a metadata-less
object storage service is desirable, the different semantics from traditional
file systems requires some considerations. In this case, should these services
get out of sync, three particular risks are introduced.  First, a client could
create a dangling entry in the metadata service that does not correspond to any
objects in the object store. Second, a client could create orphaned objects
that have no associated metadata entries. Third, updates to the metadata or
object store service should be an atomic operation, but due to a lack of
coordination, a window where the system is inconsistent appears.

Ultimately, the consistency semantics required must be determined. If a
metadata service is required and it must be in sync with the object storage
service, then additional work must be performed. In traditional file systems,
the metadata and object storage updates are atomic. With decoupling metadata
from object storage, should this atomicity still be desired, it requires both
the ability for the services to participate in a task that is part of a larger
atomic operation and a higher-level mechanism to manage the atomic operation.

Overall, while additional work is required to maintain a client-side only
metadata service, it eliminates any potential bottlenecks related to updating
metadata related to the object storage. The burden of tracking striping and
other metadata that has traditionally been part of the metadata associated with
the file system will have to be maintained by the object storage service. The
lack of a centralized, serialized bottleneck to store that information improves
concurrency.

\subsection{Comparison to \DDTns}
The \DDT project~\cite{lofstead:2012:txn} sought to develop an efficient
approach for handling ACID-style transactions in an environment with parallel
clients and multiple servers (doubly distributed). Rather than being aimed
solely at data movement operations, \DDT seeks to address the general problem
of managing any operation with multiple clients and servers.  Consider the
management of the analysis/visualization area, potentially similar to the IOD
concept. The transaction protocol is used to help manage resizing of the
resource allocation to the various analysis and visualization components.  For
the purposes of this discussion, \DDT could also be used to manage changing how
IOD processes and/or nodes are used without exposing these changes to the
client processes prematurely.  This has been described and analyzed
previously~\cite{dayal:2013:io-containers}.

The example metadata and data storage services created as part of the \DDT
project have no dependencies between transactions that prevent visibility
should an older version be incomplete. This additional, intentional requirement
by IOD offers different functionality than \DDTns's example services. In the
case of \DDTns, the functionality is more minimal, but also avoids some of the
concerns outlined below.

The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
scalability issues and demonstrated a scalable client-side coordination model
with excellent performance. The performance measured for a complex transaction
with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}.

\begin{figure}[ht]
%\vspace{-0.15in}
\centering
\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/performance}
\vspace{-0.15in}
\caption{Total Transaction Overhead}
\label{fig:performance}
%\vspace{-0.15in}
\end{figure}

At a high level, both \DDT and the IOD transactions have the same design. In
both cases, a hierarchical model is employed. In the case of \DDTns, it is a
purely client-side tree using semi-synchronous messaging. The messaging itself,
in the current implementation, uses asynchronous MPI messages. The synchronous
component comes from the timeout mechanism used to detect faults.  It forces a
level of coordination and synchronization for the protocol. For IOD, it is a
server-side tree and fully asynchronous relying on the existing Lustre fault
detection mechanism for failure detection. In both cases, there is a master in
charge of managing the transaction and a collection of workers that aggregate
into the master through second-level leaders. Beyond that, there are some
significant differences. Some of the different choices made by IOD raise some
possible concern.

The multi-leader model introduces the possibility of forcing a rollback of an
entire transaction when a partial retry might be sufficient for success. Since
the transactions are managed at a high level rather than the individual
tasks, a failure in a limited distributed task can cause the entire transaction
to fail. For example, consider 10 processes each have 5 tasks, but 3 of those
10 have an additional shared task to complete. If the task shared by the 3
processes fails on any of the three, the entire transaction would roll back
because it is a coarse-grained success/failure. If a concept like
sub-transactions at a task granularity were used, then it would be possible for
the one process that failed to report just that failure. Then the transaction
manager could reassign the resources for these 3 processes and try just that
operation again. If it now succeeds, then the overall transaction can be marked
successful only redoing the minimum amount of work required.

\begin{figure}[ht]
\centering
\includegraphics[keepaspectratio=true, width=0.5\columnwidth]{images/optimized-protocol}
\vspace{-0.15in}
\caption{Optimized Protocol}
\label{fig:optimized-protocol}
%\vspace{-0.25in}
\end{figure}

\DDT has addressed theses issues in a couple of ways. First, the
sub-coordinators each have a list of processes from which they expect messages.
Should a message be missed, it is noticed and corrective action can be taken.
Second, \DDT has the concept of sub-transactions. The messaging requirements
are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
represent finer grained operations than the entire output, \DDT can manage
multiple writes per client by using a sub-transaction to represent the output
for any item to the file (container). Because of how the sub-transactions are
managed, the singleton sub-transactions, ones in which only a single process
participates, must be declared before the transaction begins so that its
existence can be broadcast as part of the begin transaction message. This
ensures there is global knowledge that the sub-transaction is expected.  That
way if the coordinator (transaction leader) fails, whichever process takes over
that role knows to expect a completion message for that sub-transaction or the
overall transaction cannot complete. Global sub-transactions can be defined at
any time since they are a global, synchronized operation broadcasting their
existence. While this additional layer does introduce messaging, the overhead
is quite small.

The advantages of eliminating these messages is not performance as demonstrated
by the performance of \DDTns. Instead, it offers a much less synchronous model
that matches with different programming models, such as Charm++ or other
task-based approaches. Since it can work for the bulk-synchronous model also,
it is a more broadly applicable approach. This assumes that the observed
potential issues can be addressed successfully.

@@

>>>>

The rest of the paper is organized as follows. A brief overview of related work
is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array. This will be discussed in the contect of the HDF-5 based
example library used for the functionality demonstration. Section~\ref{sec:iof}
briefly discusses the motivation and proposal for the IO forwarding layer.
Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
functionality it offers. This will detail the pieces of the layer that are
potentially optional and mention the cross-cutting features discussed in a
later, cross-cutting section. Section~\ref{sec:daos} discusses how the
persistent storage array interface itself functions. As with the IOD layer, the
disucssion of cross-cutting features will be mentioned, but disucssed more
fully in the cross-cutting section. The VOSD layer is disucssed in
Section~\ref{sec:vosd}. In particular, the mapping between the DAOS and VOSD
layers are explored as it pertains to the physical storage. Next is an
exploration of cross-cutting features like transactions and metadata management
in Section~\ref{sec:broader}. Since these and other features are spread across
multiple layers, it makes more sense to discuss them independently once an
understanding of the overall structure has been presented.  A demonstration
of the functionality is presented in Section~\ref{sec:evaluation}. This shows
that the prototype functionality validation prototype can function. The paper
is concluded in Section~\ref{sec:conclusion} with a summary of the broad issues
covered in the paper.

@@

The IOD VOL plugin serves as the bridge between HDF5 and the IOD API (Figure \ref{fig:vol-arch}). The application calls the HDF5
library while running on the system's compute nodes. Using the VOL
architecture, the IOD VOL plugin uses a function shipper (RPC library)
to forward the VOL calls to a server component running on the I/O nodes
(IONs). At that point, the VOL calls are translated into I/O Dispatcher
(IOD) API calls and executed at the IONs.

@@

The IO Forwarding layer offers a mechanism to reduce the concurrency impact of
the massive process count on the storage stack. A current trend mixing MPI with
a threading library like OpenMP is addressing the same issue. In this case,
the storage stack has far fewer end-points in which to receive requests and
data from compute processes. By reducing the number of simultaneous requests,
delays can be reduced. This has been demonstrated for the file open operation
with Lustre~\cite{lofstead:2009:adaptable} and to some degree for accessing the
storage devices themselves~\cite{lofstead:2010:io-variability}. The BlueGene
platform incorporated dedicated hardware to perform this role. The proposed
functionality for this layer has no special features beyond those previously
demonstrated.

@@

%In current proposals, the stack is managed by middleware that sits between the
%application and the parallel file system. Since most of the applications assume
%a POSIX interface to storage, existing middleware either transparently handle
%the I/O operations, appearing as regular POSIX calls to applications; or modify
%the I/O API as little as possible in order to minimize the impact on production
%codes.
%
%As we move towards the exascale goal, this way of interfacing with storage
%becomes more of an obstacle. Many features provided by the stack are hidden by
%the middleware/POSIX layers. If these were visible to the applications, and if
%the applications were able to control how the underlying layers behave, they
%would benefit greatly since the domain knowledge that is available at the
%application level could be used to execute storage operations efficiently. One
%more implication of the 25-year-old POSIX API is in cumbersome middleware,
%since metadata has to be part of the same uni-dimensional byte array being
%stored.
%
%The Fast Forward Storage I/O project is aimed at merging the features of
%existing middleware into a next-generation parallel file system stack. This new
%storage API termed the I/O dispatcher (IOD) is object-based, transactional,
%asynchronous and I/O-staging-aware (Figure~\ref{fig:exa-stack} (right)).

@@

%The prototype stack, from a top-down point of view, has the following
%components:
%
%\begin{enumerate}
%\def\labelenumi{\arabic{enumi}.}
%\itemsep1pt\parskip0pt\parsep0pt
%\item
%  Arbitrarily Connected Graphs (ACG).
%\item
%  HDF5 extensions.
%\item
%  I/O Dispatcher (IOD).
%\item
%  Distributed Asynchronous Object Storage (DAOS).
%\item
%  Versioning Object Storage Device (VOSD).
%\end{enumerate}
%
%Applications running on the compute nodes will be written in Python scripts or
%in C++ using the HDF5 APIs. These applications will execute analysis on
%graph-based data (ACG). The point is to demonstrate both the HPC and BigData
%use cases of the Exascale architecture. The data structures will be stored in
%HDF5, for which new API extensions will be implemented. These new API calls
%will expose the transactional, asynchronous and function-forwarding semantics
%of the underlying stack.

@@

@@

%One of the bigger concerns is the observation that the original data staging
%proposals all used compute nodes while the newer proposals seek not only to
%make them a fixed portion of the IO stack, but also shared across all machine
%users. The PreDatA~\cite{zheng:2010:predata} paper in particular examines the
%potential costs and advantages of where to place operators similar to the IOD
%proposed function shipping. There are two key takeaways from PreDatA. First,
%placement matters.  Depending on the communication intensity vs. computation
%intensity, where along the IO path to place the operation can matter
%significantly. Second, and more importantly, the amount of time spent
%processing for the operators was stretched to the point where it consumed
%nearly all of the time between IO operations. The given ratios of compute
%processes to staging process examined is representative for future extreme
%scale platforms. If anything, the ratios offer more staging processes than IOD
%processes would be available.
%
In the case of the written IOD design, it describes a fixed-sized staging area
that is partitioned on a per-application basis.
%This is unlikely to be useful
%because of the limited compute and communication capacity to spare to perform
%these operations at a bottleneck in the IO path.  The use of a separate data
%staging area intentionally separate from the IO path allows using operators on
%limited resources leaving the IO path clear for strictly data movement. A
%nuance of this design is discussed in the Design Philosophy below.
%
%By concentrating the Burst Buffers and function shipping into the storage
%stack, three problems arise.  First, the amount of network bandwidth, IO
%bandwidth, and compute power consumed for example operations from a single
%application is likely to completely monopolize the IOD processes. Second, if
%space and time partitioning is used instead, the functionality risks being too
%small to be useful. Third, the long-term hardware performance advantage for
%SSDs is questionable. Recent studies have shown that the erase-before-write
%and interference between reading and writing with flash-based SSDs can cause
%severe performance problems~\cite{skourtis:2013:ssd-performance}. The inclusion
%of an optional use SSD layer in the new Trinity machine at Los Alamos will
%offer a test bed to determine how likely these observed problems would affect a
%production extreme scale platform.
%
%Given these features, the optionality and even incorporation of burst buffers
%in the current design should be carefully considered. Much of the advanced, key
%functionality proposed as they are currently designed ultimately relies on the
%existence of burst buffers to work. Further thought about how to have an IOD
%layer both with and without a burst buffer is required before they can be
%considered optional. As the design stands today, they are a required part of
%the IOD layer for proper functioning. Unfortunately, it is not clear that they
%can address the performance concerns they are intended to cover.

@@

\section{DAOS Layer}
\label{sec:daos}

The Distributed Asynchronous Object Storage layer serves as the traditional
parallel file-system interface layer for the storage devices. This is the
consistent, global view of the underlying devices represented in this stack
by the VOSD layer.

This is the layer where the container/object model is translated into the
physical storage requirements dictated by the physical storage underneath (the
VOSD layer). The two key design elements of this layer are the handling of
epochs and the mapping of conatiners and objects to the underlying storage.

Epochs are represented throguh the shard mechanism. Each container shard is
versioned according to the transactions in the IOD layer that are persisted to
the DAOS layer as an epoch. The container shard is a piece of the container
with the contained objects concept. Since the container is largely logical,
this generally maps to a shard of an object within a container.

Each container is represented by a directory on some storage device containing
symbolic links to all of the shards it contains and maintains the epoch ID.
In particular the Highest Committed Epoch (HCE) is an important concept for
quickly identifying which version of a shard to retrieve and to block writes to
older epochs since those have been committed.

\section{VOSD Layer}\label{sec:vosd}
The Versioning Object Storage Device (OSD) layer operates as the layer for each
persistent storage device used to support the parallel storage array. In the
purest form, it uses a local file system to arrange storage of objects that
represent parts of the higher level objects in containers.

The base level implementation continues the space optimization of only storing
changes for new versions by using a copy-on-write file system. The prototype
uses ZFS~\cite{zhang:2010:zfs} for the known stability and integration with
Lustre. In a production version of the FFSIO stack,
btrfs~\cite{rodeh:2013:btrfs}, The Linux B-Tree File System, given its
open-source backing and GPL licensing, is a likely long-term choice.

At a more detailed level, the design for VOSD is an increment beyond the
current Lustre Object Storage Device design to incorporate the idea of shards
and the versioning aspects of transactions/epochs. For every DAOS shard, the
VOSD has information for storing and accessing the currently commited version,
the Highest Committed Epoch, as well as a staging dataset represeting the
next version of the object being stored. Both of these are combined in a shard
root.

For data integrity, an intent log is maiained as part of the underlying file
system enabling fault recovery.

Beyond the functionality to incorporate and expose the copy-on-write nature
of the underlying file system and the semantics for storing and processing
shards and their associated epochs, this is largely an evolution of the
existing Lustre OSD layer.

@@

%At a broader level, there are some concerns that were partially clarified
%through conversations with the team.  Consider a shared file system across an
%HPC data center. The current design maintains the metadata in its own
%container. Since copying data from the IOD layer to the DAOS layer requires an
%explicit persist call, how and when synchronizing the metadata across the
%layers and potentially across machines occurs is undefined. Delaying
%synchronization until an explicit persist is called will reduce the update
%frequency, but delays the data visibility on other platforms. Ideally, the
%metadata object would need to be automatically persisted every time a container
%transaction is persisted to the DAOS layer.
%
%The implication of this is that every transaction persist is double operation
%to account for the metadata persist. More importantly, the IOD-layer version of
%the metadata container may contain readable transactions that have not been
%persisted to the DAOS layer. How to handle this inconsistency between the two
%layers still needs to be explored.
%
A point of confusion rather than a potential design challenge is the change in
definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
container is a collection of objects. For the DAOS layer, a container is a
collection of objects across a set of shards. For the IOD an object may be a
shard of a global array.  For DAOS, a shard can host a set of DAOS objects.
Having the same names with locally correct, globally conflicting definitions
serves to confuse how the system should work.

@@

The transaction mechanism manifests in two forms. At the IOD layer, they are
called transactions and are used to judge whether or not a set of distributed,
asynchronous modifications across a set of related objects is complete or not.
It is also used to control access by treating the transaction ID of committed
transaction as a version identifier.  At the DAOS layer, they are called epochs
and represent persisted (durable) transactions from the IOD layer. Each of
these offers different functionality, but are connected as is explained below.
%How these differ from the \DDT approach is also explored.  While IOD's and
%\DDTns's transactions are seemingly very different, they use a similar
%high-level design, but very different implementation, to solve the same
%problem.

@@

%Based on the lessons from the \DDT metadata
%service~\cite{lofstead:2012:txn-metadata} construction and the prior
%experiments with LWFS, having a completely separate metadata service is
%workable. Rather than making it a bottleneck in the IO path, it is another
%service that users must interact with if they need those services.  Users can
%manage everything by maintaining the metadata including the list of objects
%themselves. However, there are drawbacks to this 100\% client-side approach.
%
%With a client-side only approach, there is a serious risk of the metadata
%service and the object store getting out of sync.  While having a metadata-less
%object storage service is desirable, the different semantics from traditional
%file systems requires some considerations. In this case, should these services
%get out of sync, three particular risks are introduced.  First, a client could
%create a dangling entry in the metadata service that does not correspond to any
%objects in the object store. Second, a client could create orphaned objects
%that have no associated metadata entries. Third, updates to the metadata or
%object store service should be an atomic operation, but due to a lack of
%coordination, a window where the system is inconsistent appears.
%
%Ultimately, the consistency semantics required must be determined. If a
%metadata service is required and it must be in sync with the object storage
%service, then additional work must be performed. In traditional file systems,
%the metadata and object storage updates are atomic. With decoupling metadata
%from object storage, should this atomicity still be desired, it requires both
%the ability for the services to participate in a task that is part of a larger
%atomic operation and a higher-level mechanism to manage the atomic operation.
%
%Overall, while additional work is required to maintain a client-side only
%metadata service, it eliminates any potential bottlenecks related to updating
%metadata related to the object storage. The burden of tracking striping and
%other metadata that has traditionally been part of the metadata associated with
%the file system will have to be maintained by the object storage service. The
%lack of a centralized, serialized bottleneck to store that information improves
%concurrency.

%\subsection{Comparison to \DDTns}
%The \DDT project~\cite{lofstead:2012:txn} sought to develop an efficient
%approach for handling ACID-style transactions in an environment with parallel
%clients and multiple servers (doubly distributed). Rather than being aimed
%solely at data movement operations, \DDT seeks to address the general problem
%of managing any operation with multiple clients and servers.  Consider the
%management of the analysis/visualization area, potentially similar to the IOD
%concept. The transaction protocol is used to help manage resizing of the
%resource allocation to the various analysis and visualization components.  For
%the purposes of this discussion, \DDT could also be used to manage changing how
%IOD processes and/or nodes are used without exposing these changes to the
%client processes prematurely.  This has been described and analyzed
%previously~\cite{dayal:2013:io-containers}.
%
%The example metadata and data storage services created as part of the \DDT
%project have no dependencies between transactions that prevent visibility
%should an older version be incomplete. This additional, intentional requirement
%by IOD offers different functionality than \DDTns's example services. In the
%case of \DDTns, the functionality is more minimal, but also avoids some of the
%concerns outlined below.
%
%The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
%scalability issues and demonstrated a scalable client-side coordination model
%with excellent performance. The performance measured for a complex transaction
%with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
%explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}.
%
%\begin{figure}[ht]
%%\vspace{-0.15in}
%\centering
%\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/performance}
%\vspace{-0.15in}
%\caption{Total Transaction Overhead}
%\label{fig:performance}
%%\vspace{-0.15in}
%\end{figure}
%
%At a high level, both \DDT and the IOD transactions have the same design. In
%both cases, a hierarchical model is employed. In the case of \DDTns, it is a
%purely client-side tree using semi-synchronous messaging. The messaging itself,
%in the current implementation, uses asynchronous MPI messages. The synchronous
%component comes from the timeout mechanism used to detect faults.  It forces a
%level of coordination and synchronization for the protocol. For IOD, it is a
%server-side tree and fully asynchronous relying on the existing Lustre fault
%detection mechanism for failure detection. In both cases, there is a master in
%charge of managing the transaction and a collection of workers that aggregate
%into the master through second-level leaders. Beyond that, there are some
%significant differences. Some of the different choices made by IOD raise some
%possible concern.
%
%The multi-leader model introduces the possibility of forcing a rollback of an
%entire transaction when a partial retry might be sufficient for success. Since
%the transactions are managed at a high level rather than the individual
%tasks, a failure in a limited distributed task can cause the entire transaction
%to fail. For example, consider 10 processes each have 5 tasks, but 3 of those
%10 have an additional shared task to complete. If the task shared by the 3
%processes fails on any of the three, the entire transaction would roll back
%because it is a coarse-grained success/failure. If a concept like
%sub-transactions at a task granularity were used, then it would be possible for
%the one process that failed to report just that failure. Then the transaction
%manager could reassign the resources for these 3 processes and try just that
%operation again. If it now succeeds, then the overall transaction can be marked
%successful only redoing the minimum amount of work required.
%
%\begin{figure}[ht]
%\centering
%\includegraphics[keepaspectratio=true, width=0.5\columnwidth]{images/optimized-protocol}
%\vspace{-0.15in}
%\caption{Optimized Protocol}
%\label{fig:optimized-protocol}
%%\vspace{-0.25in}
%\end{figure}
%
%\DDT has addressed theses issues in a couple of ways. First, the
%sub-coordinators each have a list of processes from which they expect messages.
%Should a message be missed, it is noticed and corrective action can be taken.
%Second, \DDT has the concept of sub-transactions. The messaging requirements
%are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
%represent finer grained operations than the entire output, \DDT can manage
%multiple writes per client by using a sub-transaction to represent the output
%for any item to the file (container). Because of how the sub-transactions are
%managed, the singleton sub-transactions, ones in which only a single process
%participates, must be declared before the transaction begins so that its
%existence can be broadcast as part of the begin transaction message. This
%ensures there is global knowledge that the sub-transaction is expected.  That
%way if the coordinator (transaction leader) fails, whichever process takes over
%that role knows to expect a completion message for that sub-transaction or the
%overall transaction cannot complete. Global sub-transactions can be defined at
%any time since they are a global, synchronized operation broadcasting their
%existence. While this additional layer does introduce messaging, the overhead
%is quite small.
%
%The advantages of eliminating these messages is not performance as demonstrated
%by the performance of \DDTns. Instead, it offers a much less synchronous model
%that matches with different programming models, such as Charm++ or other
%task-based approaches. Since it can work for the bulk-synchronous model also,
%it is a more broadly applicable approach. This assumes that the observed
%potential issues can be addressed successfully.

@@

\section{Demonstration}
This stack has a prototype implementation intended to test concepts rather
than performance and scalability. It has focused on examining the interaction
of the different APIs for each layer to flesh out any detailed requirements
or concerns that may have been missed in the conceptualization of this IO stack.

We run this demonstration using two different modes. A standard HDF-5 test
application is used with the native file system layer writing directly to
storage is measured. Then, the same test is run against the requirements
prototype to demonstrate that it can function for this example as well. The
performance of both of these tests are reported to give a very rough idea of
the overhead that might be involved. Rather than a true overhead, this should
be considered the maximum overhead that should be expected once an optimized,
fully functional IO stack is deployed without relying on translating to an
underlying parallel file system. Figure~\ref{fig:evaluation} shows the
relative performance for both test runs.

The testing evironment for this demonstration is ....

\begin{figure}[htbp]
\centering
\includegraphics[width=\columnwidth]{images/graphlab-and-graphbuilder.png}
\caption{Functionality Demonstration Validation}
\label{fig:evaluation}
\end{figure}
